Artificial Ecosystem Selection: Methane Oxidation

### 2020-07-28 Manuscript/manuscript.Rmd
add p6

dev_fit <- rma(deviance~passage, se^2, method="FE", data = deviance)
dev_fit2 <- lm(deviance~passage, data = deviance)

@@

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
  geom_pointrange(color = 'darkorange2') + 
  #geom_abline(intercept = 0.1 + fit$beta[1], slope = fit$beta[2], color = 'blue') +
      geom_segment(aes(x = 1, xend = 5, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*5), color = 'darkorange2', size = 1) +
  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
  theme_bw()
```

@@

Methane oxidation rate was heritable between consecutive passages (`r figures("herit", display="cite")`). Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)`. With an inbreeding coefficient of `r myround(F, 2)`,

>>>>

dev_fit <- rma(deviance~passage, se^2, method="FE", data = deviance)

@@

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
  geom_pointrange(color = 'darkorange2') + 
      geom_segment(aes(x = 1, xend = 6, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*5), color = 'darkorange2', size = 1) +
  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
  theme_bw()

@@

Methane oxidation rate was heritable between consecutive passages (`r figures("herit", display="cite")`). Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)` (t = `r myround(her_sum[2, 4], 2)`, p = `r myround(her_sum[2, 5], 3)`). With an inbreeding coefficient of `r myround(F, 2)`,

### 2020-07-28 R/process.Rmd
add p6

### Standard Curve

```{r sc_data, fig.height=3}

@@

sc_ch4 <- 
sc_ch4 %>% 
  left_join(sc_dates) %>% 
  select(flux_date:sc_date, t, r.squared:slope)

@@

### Flux Measurements

@@

# Flux

@@

flux_data %>% 
  ggplot(aes(days, ppm)) + 
  facet_grid(flux_date + treat ~ rep, scales = "free") +
  geom_point() + 
  stat_smooth(method = 'lm')

@@

@@

```{r test_response}

@@

```{r selection}

@@

```{r selection_log_plot}

@@

fit <- lm(estimate ~ treat * passage, data = log_fluxes)
summary(fit)
plot(fit$residuals ~ log_fluxes$estimate)

@@

```{r test_response_rank}

@@

```{r}
selection_table <- tidy(anova(rank_fit))
selection_table[1, 1] <- "Treatment"
selection_table[2, 1] <- "Passage"
selection_table[3, 1] <- "Treatment:Passage"
selection_table[4, 1] <- "Residuals"
selection_table %>% 
  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
  kable_styling()

@@

```{r selection_rank}

@@

gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

ggplot(fluxes, aes(x = estimate)) + 
  facet_wrap(~ passage + treat) +
  geom_dotplot(method = 'histodot') +
  geom_vline(aes(xintercept = mean(estimate)), color = 'blue') +
  geom_vline(aes(xintercept = median(estimate)), color = 'red') +
    geom_vline(aes(xintercept = gm_mean(estimate)), color = 'green')

@@

```{r}
deviance_se <- 
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(se = se(estimate)) %>% 
  pivot_wider(names_from = treat, values_from = se) %>% 
  mutate(deviance_se = p + n) %>% 
  pull(deviance_se)

deviance <-
fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate)) %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

@@

fit <- lm(deviance ~ passage, data = deviance)
summary(fit)

fit <- rma(deviance~passage,se^2,method="FE",data = deviance)
fit

summary(fit)
plot(fit)

@@

@@

parental <-
  parental %>% 
  group_by(passage, treat) %>% 
  summarize(parental = mean(estimate)) %>% 
  ungroup() %>% 
  filter(passage != 5)

@@

offspring <-
  offspring %>% 
  group_by(passage, treat) %>%
  summarize(offspring = mean(estimate)) %>%
  ungroup() %>%
  # mutate(offspring = estimate) %>%
  filter(passage != 1) %>% 
  mutate(passage = passage - 1)

@@

ggsave('heritability.pdf', width = 4, height = 3)
ggsave('heritability.png', width = 4, height = 3)
heritability$parental_sq <- heritability$parental^2
fit <- lm(sqrt(offspring) ~ parental, data = heritability)

>>>>

# Standard Curves

Standard curves for each sc_date. (Some flux_dates have multiple standard curves if the instrument drifted from day to day as indicated by check standards.)

```{r sc_data, fig.height=4}

@@

sc_ch4 <- 
sc_ch4 %>% 
  left_join(sc_dates, by = "sc_date") %>% 
  select(flux_date:sc_date, t, r.squared:slope)

@@

Each column is jar by number (the jar numbers are arbitrary - there is no relationship between the communities of jar 01, passage 1 and jar 01, passage 2). Rows are each selection treatment (positive = p, neutral = n) within each passage (indicated by the date flux was measured). Lines are proportion of ppm CH4 consumed per day.

@@

@@

flux_data %>% 
  ggplot(aes(days, ppm)) + 
  facet_grid(flux_date + treat ~ rep, scales = "free") +
  geom_point() + 
  stat_smooth(method = 'lm', formula = y ~ x)

@@

Histograms and boxplots of flux estimates for each jar within each treatment. The numbered points give an indication of which jars were the greatest methane consumers

@@

# Response to Selection

```{r test_response, eval=FALSE, include=FALSE}

@@

```{r selection, eval=FALSE, include=FALSE}

@@

```{r selection_log_plot, eval=FALSE, include=FALSE}

@@

fit <- lm(estimate ~ treat * passage, data = log_fluxes)
plot(fit$residuals ~ log_fluxes$estimate)

@@

```{r test_response_rank, eval=FALSE, include=FALSE}

@@

```{r eval=FALSE, include=FALSE}
selection_table <- tidy(anova(rank_fit))
selection_table[1, 1] <- "Treatment"
selection_table[2, 1] <- "Passage"
selection_table[3, 1] <- "Treatment:Passage"
selection_table[4, 1] <- "Residuals"
selection_table %>% 
  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
  kable_styling()

@@

```{r selection_rank, eval=FALSE, include=FALSE}

@@

@@

```{r}
deviance_se <- 
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(se = se(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = se) %>% 
  mutate(deviance_se = p + n) %>% 
  pull(deviance_se)

deviance <-
fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

@@

fit <- rma(deviance~passage,se^2,method="FE",data = deviance)

@@

rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)
rma_output %>% 
  kable() %>% 
  kable_styling()
```

@@

parental <-
  parental %>% 
  group_by(passage, treat) %>% 
  summarize(parental = mean(estimate), .groups = "drop")

@@

offspring <-
  offspring %>% 
  group_by(passage, treat) %>%
  summarize(offspring = mean(estimate), .groups = "drop") %>%
  # mutate(offspring = estimate) %>%
  filter(passage != 1) %>% 
  mutate(passage = passage - 1)

@@

ggsave('heritability.pdf', width = 4, height = 3)
ggsave('heritability.png', width = 4, height = 3)

heritability$parental_sq <- heritability$parental^2
fit <- lm(sqrt(offspring) ~ parental, data = heritability)
